---
title: "Project_CIS603"
author: "Saikiran Jelloji"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(randomForest)
library(datasets)
library(caret)
```

## Summary

The project has included more of visualization to understand the distribution of main air pollutants, sulphur and carbon, in urban areas. The project has considered different cities and main air pollutants for each city. The project has included a multiple linear regression model that can be used for predicting the occurence of air pollution from the main air pollutants like the sulphur and carbon. The project has illustrated that sulphur is the leading air pollutant in the urban areas. The p-value of the linear regression model has illustrated that the SO2.Mean, CO.Mean, and NO2.Mean are not  statistically significant to our model. 

## Introduction

Air pollution has become a serious challenge in urban areas. This is because air pollution is associated with diverse health issues like cancer, a deadly disease, or heart disease. Due to the high number of individuals, industries, and traffic, urban areas have been identified as the most vulnerable to air pollution. Therefore, my project tries to identify the factors that are highly causing air pollution in urban areas and take the right actions. Therefore, the importance of this project provides the advantage of identifying the main causes of air pollution and also forecast the levels of air pollution in urban areas. With the identification and forecasting, it will be helpful to create awareness among people on taking precautions that would be vital to avoid being exposed to air pollutants. In addition, the project is important to the environmental regulatory boards in implementing regulations that might help reduce air pollution in urban areas. 

## Literature Review

There have been research papers that have shown interest in the predictions of the levels of air pollution by the use of techniques associated with statistical models. The research paper by Mustafaraj et al. (2011) has implemented a technique that used both auto regressive and neural network models to predict air pollution levels. Another study by Qin & Gong (2022) incorporated the use of machine learning methods for the predictions of Carbon dioxide levels in China. 

## Research Question

The research question for the project is:

1.What is the data analysis method or technique applicable in R for predicting the levels of air pollution?

## Data
The data used in the project has been obtained from sites that store data associated with environmental factors. The data contains the nitrogen and carbon dioxide levels variables that are measured in the air and information on the temperature, humidity, or wind speed. 

```{r echo=FALSE}
data <- read.csv("C:/Users/saiki/Downloads/R project/air_pollution_us_urban.csv")
head(data)
```
Check for if there are any missing values in the data. 
```{r echo=FALSE}
colSums(is.na(data))
```
## Visualizations

## Histogram for Mean Nitrogen Levels
```{r echo=FALSE}
##convert NO2.Mean variable to numeric
data$NO2.Mean <- as.numeric(data$NO2.Mean)
##Create a histogram for NO2.Mean
Nitrogen_Levels <- data$NO2.Mean
hist(Nitrogen_Levels)
```

## Histogram for Mean Suphur Levels

```{r echo=FALSE}
##convert SO2.Mean variable to numeric
data$SO2.Mean <- as.numeric(data$SO2.Mean)
##Create a histogram for SO2.Mean
Sulphur_Levels <- data$SO2.Mean
hist(Sulphur_Levels)
```

## Histogram for Carbon Levels

```{r echo=FALSE}
##convert CO.Mean variable to numeric
data$CO.Mean <- as.numeric(data$CO.Mean)
##Create a histogram for SO2.Mean
Carbon_Levels <- data$CO.Mean
hist(Carbon_Levels)
```

## Barchart fot Nitrogen Levels in Cities

```{r echo=FALSE}
# Sort NO2.Mean (Nitrogen LEvels) in decreasing order
sort_data<-data[order(data$NO2.Mean, decreasing = TRUE), ] 
```
```{r echo=FALSE}
# Subset data frame to first 10 cities
nitrogen_10 <- head(sort_data, 10)

# Create plot
ggplot(nitrogen_10, aes(x = City, y = CO.Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  xlab("City") +
  ylab("Nitrogen Level") +
  ggtitle("Nitrogen Levels Cities")
```

## Barchart fot Carbon Levels in Cities

```{r echo=FALSE}
# Sort C0.Mean (Carbon Levels) in decreasing order
sort_data<-data[order(data$CO.Mean, decreasing = TRUE), ] 
```
```{r echo=FALSE}
# Subset data frame to first 10 cities with highest carbon levels
carbon_10 <- head(sort_data, 10)

# Create plot
ggplot(carbon_10, aes(x = City, y = CO.Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  xlab("City") +
  ylab("Carbon Level") +
  ggtitle("Carbon Levels in Cities")
```

## Barchart fot Sulphur Levels in Cities

```{r echo=FALSE}
# Sort SO2.Mean (Sulphur Levels) in decreasing order
sort_data<-data[order(data$SO2.Mean, decreasing = TRUE), ] 
```
```{r echo=FALSE}
# Subset data frame to first 10 cities with highest carbon levels
sulphur_10 <- head(sort_data, 10)

# Create plot
ggplot(sulphur_10, aes(x = City, y = SO2.Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  xlab("City") +
  ylab("Sulphur Level") +
  ggtitle("Sulphur Levels in Cities")
```
## Analysis

Split Dataset to 80% training set and 20% testing set. 
```{r echo=FALSE}
set.seed(100)
data_splitted<- sample(2, nrow(data), replace = TRUE, prob = c(0.8, 0.2))
# Training Set Data
train_data <- data[data_splitted==1,]

# Testing Set Data
test_data <- data[-data_splitted==2,]
table(data_splitted)
```
The training set has 5593 observations and testing set has 1407 observations. 
## Build the predicting Linear Regression model

```{r echo=FALSE}
#fit model
model <- lm(Respiratory.Diseases~ SO2.Mean+CO.Mean + NO2.Mean, data=train_data)

#view model summary
summary(model)
```
The equation of the linear regression model for predicting the probability of respiratory diseases through the three main air pollutants, sulphur, carbon, and nitrogen can be formulated as:
Predicted= 0.086+0.0006991* SO2.Mean- 0.01542* CO.Mean+0.007094*NO2.Mean. 
For example:
Predicted= 0.086+0.0006991* 10- 0.01542* 20+0.007094*15
If answer is:
1 represents yes there is respiratory disease for example lung cancer
0 represents no respiratory disease .

## Conclusion

The project has been successfully performed and it has been clear on the main air pollutants in urban areas and in different cities. Sulphur is main air pollutant in urban areas. Also the linear regression model have shown that the predictor variables are not  statistically significant to our model (Maverick, 2023). 

## References
Maverick. (2023). Learn to Predict Using Linear Regression in R With Ease (Updated 2023). Analytics Vidhya.               https://www.analyticsvidhya.com/blog/2020/12/predicting-using-linear-regression-in-r/
Mustafaraj, G., Lowry, G., & Chen, J. (2011). Prediction of room temperature and relative humidity by autoregressive linear and nonlinear neural network models for an open office. Energy and Buildings, 43(6), 1452-1460.
Qin, J., & Gong, N. (2022). The estimation of the carbon dioxide emission and driving factors in China based on machine learning methods. Sustainable Production and Consumption, 33, 218-229.

